---
title: "P8105_Homework_1"
author: "Cecilia Zhao"
date: "2026-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains data on penguin species, island, bill length, bill depth, flipper length, body mass, sex, and year. The dataset includes three penguin species: Adelie, Chinstrap, and Gentoo, which are observed on three islands: Biscoe, Dream, and Torgersen. Also, the dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns.The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r, warning=FALSE}
ggplot(data = penguins,
       aes(x = bill_length_mm,
           y = flipper_length_mm,
           color = species)) +
  geom_point()
```

```{r, warning=FALSE}
ggsave("penguins_scatterplot.png")
```


## Problem 2

```{r}
q2_df = 
  tibble(
    rand_samp = rnorm(10),
    log_vec = rand_samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    fac_vec = factor(c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A"))
  )
```

```{r}
mean(pull(q2_df, rand_samp))

mean(pull(q2_df, log_vec))

mean(pull(q2_df, char_vec))

mean(pull(q2_df, fac_vec))
```

The mean works for the numeric and logical variables, but not for the character or factor variables.

```{r, results='hide', warning=FALSE}
as.numeric(pull(q2_df, log_vec))
as.numeric(pull(q2_df, char_vec))
as.numeric(pull(q2_df, fac_vec))
```

`as.numeric()` converts logical variable to 1s and 0s, where `TRUE` is 1 and `FALSE` is 0. The character variable is converted to `NA`. The factor variable is converted to 1, 2, and 3 respectively based on its underlying integer codes (A is 1, B is 2, and C is 3). This helps explain why `mean()` works for logical variables but not directly for character or factor variables.